Reasoning Graphically about Quantum Computation

### 2008-02-28 papers/reasoning-graphically/paper.tex
slight restructuring and some comcl stuff
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1381 f8f375fb-4f10-0410-8d47-b6fbdc7ca053
inprogress on reasoning section
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1382 f8f375fb-4f10-0410-8d47-b6fbdc7ca053

\subsection{Graph Representations for Compact Closed Categories}
\label{sec:graph-repr-comp}

@@

\begin{definition}
  Two mappings $f_G$ and $f_H$ \emph{agree} when $\forall v \in
  f_G(G).\, f_H(v) \in f^{-1}_G(v)$ and $\forall v \in f_H(H).\,
  f_G(v) \in f^{-1}_H(v)$. Intuitively this captures the idea a node
  is one function is always mapped to one the nodes that maps to it in
  the other function.
\end{definition}

@@

The purpose of graph patterns is to formally express, in a declarative
manner, the intuitive idea of graph matching commonly used in for
Quantum computation. To do this, we combine the two extensions to the
concrete graph representation, namely variable nodes and !-boxes. This
provides a matching relation between graph patterns and concrete
graphs as well as directly between graph patterns. The former relation
gives a semantics for our system and ensures soundness, while the
latter supports deirved equations and thus provides the theoretical
basis for our formulation of rewriting.

@@

The semantics for a graph pattern $G$ in terms of a set of concrete
graphs, is denoted by $\minterp{G}$. We say that members of this set
are \emph{matched} by $G$. Formally, the set of matches are defined as
follows:

$\minterp{G} = (\vinterp{G'} | G' \in \binterp{G})$

@@

The definition of one pattern matching another one is simply that it
is more general. Formally, a graph pattern, $G$, matches another graph
pattern, $H$, when $\minterp{G} \supseteq \minterp{H}$. This provides
the specification for a matching algorithm which we implement of
in~\S\ref{sec:rewriting}. To make such a proof easier, we extend
embedding and plugging of variable-node graphs to graph patterns.

Our representation has been carefully designed so that this concept
corresponds to what we call a \emph{open embedding} of $G$ into
$H$.

@@

***

This is the basis for our implementation and is thus also the basis
for its correctness argument.

\section{Graph Pattern Rewriting}
\label{sec:rewriting}

\subsection{Rules}
\subsection{Unification}
\subsection{Substitution}

\section{Node Expressions}
\label{sec:node-expressions}

>>>>

A strict symmetric monoidal  category
\cite{MacLane:CatsWM:1971,AspLon:CatTypStruct:1991} is called compact
closed \cite{KelLap:comcl:1980} when each object $A$ has a chosen dual
object $A^*$, and morphisms
\begin{gather*}
  d_A : I \to A^* \otimes A\\
  e_A : A \otimes A^* \to A
\end{gather*}
such that 
\begin{align*}
  A \iso A \otimes I \rTo^{\id{A} \otimes d_A} A \otimes A^* \otimes A 
  \rTo^{e_A \otimes \id{A}} I \otimes A \iso A
\end{align*}
\subsection{Graph Representations for Compact Closed Categories}
\label{sec:graph-repr-comp}

@@

\begin{definition}
  Two mappings $f_G$ and $f_H$ \emph{agree} when $\forall v \in
  f_G(G).\, f_H(v) \in f^{-1}_G(v)$ and $\forall v \in f_H(H).\,
  f_G(v) \in f^{-1}_H(v)$. This states when a set of nodes in one
  graph all map to a common node $n$ in the other graph, $n$ maps back
  to one of the nodes that maps to it. 
\end{definition}

@@

The purpose of graph patterns is to formally express, in a declarative
manner, the intuitive idea of graph matching commonly used in for
Quantum Computation. To do this, we combine the two extensions to the
concrete graph representation, namely variable nodes and !-boxes, to
provides a matching relation between graph patterns and concrete
graphs as well as directly between graph patterns. The former relation
gives a semantics for our system and ensures soundness, while the
latter supports deirved equations and thus provides the theoretical
basis for our formulation of rewriting.

@@

The semantics for a graph pattern $G$ is a set of concrete graphs
denoted by $\minterp{G}$. We say that members of this set are
\emph{matched} by $G$. Formally, the set of matches are defined as:

$$\minterp{G} = (\vinterp{G'}\;|\; G' \in \binterp{G})$$

@@

The specification for one pattern matching another one is simply that
it is more general. Formally, a graph pattern, $G$, matches another
graph pattern, $H$, when $\minterp{G} \supseteq \minterp{H}$. This is
the specification for a matching algorithm which we implement of
in~\S\ref{sec:rewriting}. We can use the definition of open embedding
from variable-node graphs as an intermediate mechanism and show that a
suitable definition of graph pattern matching is simply the !-Box
matching followed by variable-node matching:

\begin{definition}
  $G \leq H \Leftrightarrow \forall H' \in \binterp{H}.\,\exists G' \in \binterp{G}.\, G' \leq_v H'$
\end{definition}

Because viable-node matching preserves the semantics, it is trivial to
prove that this definition meets the specification. This gives us
matching between pattern-nodes.

\subsubsection{Graph Pattern Equality.}

Equality of graph patterns can simply be defined as mutual matching:

\begin{definition}
  $G = H \Leftrightarrow G \leq H \land H \leq G$
\end{definition}

@@

\subsection{Adequacy}

\section{Reasoning with Graph Patterns}
\label{sec:rewriting}

The formalism we have presented for graph patterns provides a
\emph{meta-level} framework for reasoning in some model of concrete
graphs. The chosen model provides the \emph{object-level} graph
formalism. This includes a particular representation for data at the
nodes and edges as well as node-matching and edge-matching
algorithms. The object-level definition also defines a set of
axiomatic equational rules. These are derived

\subsubsection{Soundness.} An object-level graph system is sound if
their is some equation which can not be derived.

 The soundness
of a set of axioms in our framework

 framework
requres the model is an instance of a compact closed category, and
that the

For the model we use of qantum computation, this is finite
dimensional hilbert spaces. Because the basic equational rules come
from a model and have been proved for that model, the resulting
preserve the soundness of our graphical system.

 and the families of them represneted by graph
patterns.

An object graph formalsism uses the

By defining graph patterns as a conservative extension of concrete
graphs we get the relative soundness for them.

Such a theorem prover starts with a system of equations between
graphs. These are not true of the graphs abstractly, but are proved of
the particular model of concrete graphs being considered.

\subsection{Rules}

\subsection{Unification}
\subsection{Substitution}

\subsection{Node in Quantum Computations}
\label{ssec:node-expressions}

 Expressions
